A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-06 paper/ms.tex
update ms / Rp|P|age

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
    Color-magnitude diagram of the \cn, IC\,2602 ($\approx38$\,Myr),
    the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
    Gaia EDR3 Catalog of Nearby Stars (gray background;
    log-stretch density).  The right panel provides a zoomed view of stars on
    the pre-main-sequence.  The \cn\ and IC\,2602 are approximately
    the same isochronal age.  {\it Bottom:} TESS and Kepler stellar
    rotation periods versus dereddened Gaia colors, with the Pleiades
    and Praesepe (650\,Myr) shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.  The \cn\ is
    gyrochronally younger than the Pleiades.
   \label{fig:age}
	}
\end{figure*}

@@

To summarize, the age of the \cn\ cluster based on the empirical
procedure described in Section~\ref{sec:cmd} is
\clusterage.  This procedure hinges on the age adopted for
IC\,2602, since it is the closest reference cluster in the
color-magnitude diagram.  Gyrochronologically, while most of the G and
K dwarfs in the \cn\ have not yet converged to a slow sequence, their
rotation periods and those of the F dwarfs are also consistent with a
30 to 50\,Myr age. The most precise age for \sn\ and its
planet are therefore also \clusterage\ Myr, based on the star's
spatial, kinematic, and rotational connection to the cluster,
and the assumption that the planet formed very quickly after the star.

@@

If the \pn\ transit signal is created by a bonafide planet,  then to
our knowledge its age would make it the youngest planet yet found by
the main Kepler mission.\footnote{The re-purposed K2
mission however has found two younger systems: K2-33b and V1298 Tau
\citep{David_et_al_2017,Mann_K2_33b_2016,david_four_2019}}  A portion
of the light curve is shown with the phase-folded transit in
Figure~\ref{fig:lc}.  The dominant signal in the \texttt{PDCSAP}
photometry is a quasi-periodic starspot signal with a peak-to-peak
amplitude that varies between 2\% and 8\%, depending on the degree of
asymmetry in spot coverage between the stellar hemispheres.  The
starspot signal repeats every $2.642\pm0.042$\,days, where the
uncertainty is based on the scatter observed between different Kepler
quarters.  The details of how we model the starspot signal to derive
the parameters of the planet are discussed in
Appendix~\ref{app:gptransit}.  The stellar parameters are listed in
Table~\ref{tab:starparams}; the stellar mass, radius, and effective
temperature are found by interpolating against a 38$\,$Myr MIST
isochrone.  The systematic uncertainties are taken to be the
difference in parameters from the PARSEC isochrones, and the reported
uncertainties are a quadrature sum of the statistical and systematic
components.  The transit-fitting results, given in
Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
planet ($3.34\pm 0.18\,R_\oplus$) on a circular\footnote{Note $e<0.48$
at 2-$\sigma$; the constraints on the eccentricity are not
particularly strong.} 7.2\,day orbit, around a G8V star ($0.88 \pm
0.02 R_\odot$).

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} validated
the planet based on the transit shape, arguing that the only plausible
false positive scenario was that of a background eclipsing binary,
reported to have a model-dependent probability of $\approx10^{-5}$.
As discussed in Appendix~\ref{app:companionstar},
we discovered a low-mass stellar companion ($M_{\rm B}\approx
0.44\,M_\odot$), and so here we reconsider this false positive scenario
in some detail.

@@

Three new lines of evidence support the planetary interpretation.
First, the transit duration and the orbital period are inconsistent
with an eclipsing body around the M-dwarf companion: we find
$\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
theoretically expected density for the companion at 38\,Myr is
$\approx 3.2\,{\rm g\,cm}^{-3}$ (Table~\ref{tab:posterior};
\citealt{choi_mesa_2016}).  The transit duration is therefore too long
to explained by a star eclipsing the M dwarf secondary at
$5$-$\sigma$ confidence.

@@

The second line of confirmatory evidence for the planetary
interpretation comes by analyzing the individual transit times in the
Kepler data. We isolated each of the 144 observed transits to within
$\pm4.5$\,hr of each transit, and fitted each window with both {\it
i)} a local second-order polynomial and transit, and {\it ii)} a local
linear trend.  We let the mid-time of each transit float, and then
calculated the residual between the measured mid-time and that of a
perfectly periodic orbit.  This residual, the transit timing variation
(TTV), is plotted in Figure~\ref{fig:obliquity} against the local
linear slope.  A significant correlation of $-0.0564 \pm 0.0098\ {\rm
ppt\,day}^{-1}$ is observed.  Less than ten Kepler Objects of
Interest have shown this correlation \citep{holczer_time_2015}, which
is most readily interpreted as a TTV induced by unresolved starspot
crossings \citep{mazeh_time_2015}.  This is only possible if the
planet transits the primary star, which excludes a background
eclipsing binary scenario, and therefore confirms that \pn\ is a
planet.  It also implies that the planet's orbit is prograde.  The
latter interpretation assumes that the dominant photometric
variability is induced by dark spots, rather than bright faculae
\citep{mazeh_time_2015}.  Given the observed transition of Sun-like
stellar variability from spot to faculae-dominated regimes, we expect
this inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

@@

\begin{figure*}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f14.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radii, orbital periods, and ages of known transiting
    exoplanets}.  Planets older than $10^9$ years are shown with gray
    markers.  Planets younger than a gigayear, with $t/\sigma_t > 3$,
    are shown in color.  \pn\ is shown with a star.  The large radii
    of the youngest transiting planets could be explained by their
    primordial atmospheres not yet having evaporated; direct
    measurements of the atmospheric outflows or planetary masses would
    help to confirm this expectation.  Selection effects may also be
    important.  Parameters are from the NASA Exoplanet Archive (July
    2, 2021).
		\label{fig:rp_period_age}
	}
\end{figure*}

@@

In the context of the known transiting planets, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precisely measured ages below 100\,Myr include K2-33b,
DS~Tuc~Ab, HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet
V1298~Tau system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
\pn\ is currently the smallest of these systems ($\approx
3.3\,R_\oplus$), which may hint at selection effects in the 
young planet sample \citep{zhou_2021_tois}.  Assuming that these planets
have masses between $\approx$5 and 20$\,M_\oplus$, their large sizes
and presumed low densities are in accord with the expectation that
mini-Neptunes start their lives with large primordial atmospheres that
are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
suggestion of an aligned orbit differs from observed misalignments in number of other close-in
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Unlike these older planets, the properties of \pn\
seem to be consistent with a quiescent disk-driven migration history.
The alternative scenario of in-situ formation is disfavored on
theoretical grounds: the required disk mass would likely drive
migration, and so lacks internal consistency
\citep{inamdar_formation_2015,ogihara_reassessment_2015}.

>>>>

		\vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{f15c.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf The \cn\ is \clusterage\ old.}  {\it Top:}
    Color-magnitude diagram of the \cn, IC\,2602 ($\approx35$\,Myr),
    the Pleiades ($\approx 115$\,Myr; \citealt{dahm_2015}), and the
    Gaia EDR3 Catalog of Nearby Stars (gray background density;
    log-stretch).  The right panel provides a zoomed view of stars on
    the pre-main-sequence.  The \cn\ and IC\,2602 are approximately
    the same isochronal age.  {\it Bottom:} TESS and Kepler stellar
    rotation periods versus dereddened Gaia colors, with the Pleiades
    and Praesepe (650\,Myr) shown for reference
    \citep{rebull_rotation_2016a,douglas_poking_2017}.  The \cn\ is
    gyrochronally younger than the Pleiades.
   \label{fig:age}
	}
\end{figure*}

@@

To summarize, the age of the \cn\ cluster based on the empirical
pre-main-sequence interpolation is \clusterage.  This procedure hinges
on the age adopted for IC\,2602, since it is the closest reference
cluster in the color-magnitude diagram.  Gyrochronologically, while
most of the G and K dwarfs in the \cn\ have not yet converged to a
slow sequence, their rotation periods and those of the F dwarfs are
also consistent with a 30 to 50\,Myr age. The most precise age for
\sn\ and its planet are therefore also \clusterage\ Myr, based on the
star's spatial, kinematic, and rotational connection to the cluster,
and the assumption that the planet formed very quickly after the star.

@@

If the \pn\ transit signal is created by a bonafide planet,  then to
our knowledge its age would make it the youngest planet yet found by
the main Kepler mission.\footnote{The re-purposed K2
mission however has found two younger systems: K2-33b and V1298 Tau
\citep{David_et_al_2017,Mann_K2_33b_2016,david_four_2019}}  A portion
of the light curve is shown with the phase-folded transit in
Figure~\ref{fig:lc}.  The texttt{PDCSAP}
photometry is dominated by a quasi-periodic starspot signal with a peak-to-peak
amplitude that varies between 2\% and 8\%, depending on the degree of
asymmetry in spot coverage between the stellar hemispheres.  The
starspot signal repeats every $2.642\pm0.042$\,days, where the
uncertainty is based on the scatter observed between different Kepler
quarters.  The details of how we model the starspot signal to derive
the parameters of the planet are discussed in
Appendix~\ref{app:gptransit}.  The stellar parameters are listed in
Table~\ref{tab:starparams}; the stellar mass, radius, and effective
temperature are found by interpolating against a 38$\,$Myr MIST
isochrone.  The systematic uncertainties are taken to be the
difference in parameters from the PARSEC isochrones, and the reported
uncertainties are a quadrature sum of the statistical and systematic
components.  The transit-fitting results, given in
Table~\ref{tab:posterior}, are consistent with a mini-Neptune sized
planet ($3.34\pm 0.18\,R_\oplus$) on a close-in circular\footnote{Our
transit fitting yields $e<0.48$ at 2-$\sigma$; the constraints on the
eccentricity are not particularly strong.} orbit around a G8V star
($0.88 \pm 0.02 R_\odot$).

Could the transit signal be produced by anything other than a planet
orbiting this near-solar analog?  \citet{morton_false_2016} validated
the planet based on the transit shape, arguing that the most probable 
false positive scenario was that of a background eclipsing binary,
which had a model-dependent probability of $\approx10^{-5}$.
As discussed in Appendix~\ref{app:companionstar},
we discovered a low-mass stellar companion ($M_{\rm B}\approx
0.44\,M_\odot$), and so we reconsidered this false positive scenario
in some detail.

@@

We identified three new lines of evidence that support the planetary
interpretation.  First, the transit duration and the orbital period
are inconsistent with an eclipsing body around the M-dwarf companion:
we find $\rho_\star = 2.00 \pm 0.24\,{\rm g\,cm}^{-3}$, while the
theoretically expected density for the companion is
$\approx 3.2\,{\rm g\,cm}^{-3}$ (Table~\ref{tab:posterior};
\citealt{choi_mesa_2016}).  The transit duration is therefore too long
to explained by a star eclipsing the M dwarf secondary at
$\approx$$5$-$\sigma$.

@@

The second line of confirmatory evidence comes by analyzing the
individual Kepler transit times. We isolated each of the
144 observed transits to within $\pm4.5$\,hr of each transit, and
fitted each window with both {\it i)} a local second-order polynomial
and transit, and {\it ii)} a local linear trend.  We let the mid-time
of each transit float, and then calculated the residual between the
measured mid-time and that of a perfectly periodic orbit.  This
residual, the transit timing variation (TTV), is plotted in
Figure~\ref{fig:obliquity} against the local linear slope.  A
significant correlation of $-0.0564 \pm 0.0098\ {\rm ppt\,day}^{-1}$
is observed.  Less than ten Kepler Objects of Interest have shown this
correlation \citep{holczer_time_2015}, which is most readily
interpreted as a TTV induced by unresolved starspot crossings
\citep{mazeh_time_2015}.  This is only possible if the planet transits
the primary star, which excludes a background eclipsing binary
scenario, and therefore confirms that \pn\ is a planet.  It also
implies that the planet's orbit is prograde.  The latter
interpretation assumes that the dominant photometric variability is
induced by dark spots, rather than bright faculae
\citep{mazeh_time_2015}.  Given the observed transition of Sun-like
stellar variability from spot to faculae-dominated regimes, we expect
this inference to be reasonably secure
\citep{shapiro_are_2016,montet_long-term_2017,reinhold_stellar_2020}.

@@

\begin{figure}[tp]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{f14.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
    {\bf Radii, orbital periods, and ages of known transiting
    exoplanets}.  Planets older than $10^9$ years are shown with gray
    markers.  Planets younger than a gigayear with ${\rm
    Age}/\sigma_{\rm Age} > 3$
    are shown in color.  \pn\ is shown with a star.  The large radii
    of the youngest transiting planets could be explained by their
    primordial atmospheres not yet having evaporated; direct
    measurements of the atmospheric outflows or planetary masses would
    help to confirm this expectation.  Selection effects may also be
    important.  Parameters are from the NASA Exoplanet Archive (July
    2, 2021).
		\label{fig:rp_period_age}
	}
\end{figure}

@@

In the context of the known transiting planets, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precisely measured ages below 100\,Myr include K2-33b, DS~Tuc~Ab,
HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet V1298~Tau system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
\pn\ is currently the smallest of these systems ($\approx
3.3\,R_\oplus$), which may hint at selection effects in the young
planet sample \citep{zhou_2021_tois}.  Assuming that these planets
have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$, their
large sizes and presumed low densities are in accord with the
expectation that mini-Neptunes start their lives with large primordial
atmospheres that are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
suggestion of an aligned orbit differs from observed misalignments in
number of other close-in
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Unlike these older planets, the properties of \pn\ seem to be
consistent with a quiescent disk-driven migration history.  The
alternative scenario of in-situ formation is disfavored on theoretical
grounds: the required disk mass would likely drive migration, and so
lacks internal consistency
\citep{inamdar_formation_2015,ogihara_reassessment_2015}.
